---
title: "Introduction"
teaching: 0
exercises: 0
questions:
- "Key question (FIXME)"
objectives:
- "First learning objective. (FIXME)"
keypoints:
- "First key point. Brief Answer to questions. (FIXME)"
---

# Are you there?

Is something present in a string?

Looking at this string:

"The quick brown fox jumps over the lazy dog"

it is easy to see that the word "fox" is present.

Getting 

```{r}
library(tidyverse)

text <- "The quick brown fox jumps over the lazy dog"

str_detect(string = text, pattern = "fox")

```
That is simple. Does the book "Mathilda" contain the word mother?
```{r}
mathilda <- read_file("../data/pg15238.txt")

str_detect(mathilda, "mother")
```
It does. This is a relatively short book of only 108 pages. But it would still
take us some time to read it, and make a note of each instance of the word mother.
And given the subject, a semi-biographical tale of a young womans troubled 
relationship with her father after her mothers death, it should not come as a 
surprise

How many times does the word mother occur?

```{r}
str_count(mathilda, "mother")
```
Note the general pattern of the functions. Their names begin with *str_*, 
and continues with *what the function does*. There are two arguments. The 
text (or string), we are searching for. And the string, or pattern, that we are 
looking for.

What is the difference between a string, and a pattern?

A string is a string of characters. The way we have read in the novel, means that
the entire novel is a string. A pattern is a string that we are looking for. In
this example, the word "mother".
In order to avoid confusion, we are going to call the string we are looking for,
a pattern. The pattern we are looking for here, is the combination of the 
letters m, o, t, h, e and r. In that order. 



This motif will become very familiar.

We do have one problem. We have not actually found all the instances of 
"mother" in the text. Mother can also be capitalized, with an uppercase "M".

```{r}
str_count(mathilda, "Mother")
```

This is a common phenomenon. The standard way to handle it, is to make sure
all letters are lowercase. 

There is a function for that, and now we are invoking the pipe.
```{r}
mathilda %>% 
  str_to_lower() %>% 
  str_count("mother")
```
str_to_lower converts every letter in the text to lowercase. That makes it 
relatively easy to count the number of occurences - because we do not need to
take into account that some of the words might be capitalized.

Note that sometimes this is not a good idea. If we are trying to identify a 
character named "Miller", converting to lowercase might confuse the name with 
the profession.

str_to_lower is one member of a family of functions that changes letters from
one case to another. Let us take a look:

```{r}
test_sentence <- "this iS a senTence"
str_to_lower(test_sentence)
```
```{r}
str_to_upper(test_sentence)
```

```{r}
str_to_title(test_sentence)
```
```{r}
str_to_sentence(test_sentence)
```
Back to the novel.

A common thing to do in working with text, is to not only search, but also
replace. When we write something in word, and believe that what we are writing
about happens in London, but then find out that it is actually happening in 
Paris, it is practical to do a search-and-replace, and replace all instances
of London with Paris.

Stringr has a function for that.
str_replace
The motif is familiar. str_ at the start, and then what the function is doing,
that is replacing stuff with something else.

str_replace has a twin, str_replace_all(), that will replace all instances of
what we are looking for, whereas str_replace will replace the first instance.

The syntax is straightforward. First we provide the string we want do to 
search-and-replace on to the function. Then we give it the pattern that we want
replaced, and then the pattern we want to replace it with:

```{r}
str_replace_all("my mother lives in Helsinge", "mother", "father")
```
It can of course also be used on our example text - but it takes too long to 
read it, and note that mother has been replaced by father.

Rather than replacing a pattern with something else, it can also be used to
entirely remove that pattern:

```{r}
str_replace_all("my mother lives in Helsinge", "mother", "")
```
We do this often enough, that a function is available in stringr for this:

```{r}
str_remove_all("my mother lives in Helsinge", "mother")
```
saving a few keystrokes. The sibling to this function is str_remove() that 
will only remove the first instance of the pattern.


"Mother" is a word used in this novel.

But where in the novel? The function str_locate_all() will find the positions
in the text of the beginning and the end of the pattern we are looking for.

```{r}
falkner %>% 
  str_to_lower() %>% 
  str_locate_all("mother") %>% 
  .[[1]] %>% 
  head()
```

The functions from stringr are vectorised. That allows us to use them on
more than one text at a time. Because of this, the result is a list, in this case
with only one element. Had we applied the function to more than one text, we 
would have gotten one element in the list for each text.

the *.[[1]]* line takes what come from the pipe, indicated by the period, and
extracts the first element of that list. We are piping the result to the head()
function, in order to only see the first 6 results.

The result gives us the beginning and the end of each match in the text.

challenge
Why? mother is always six characters. If the beginning is at position 81112,
the end will always be 81117. So why would we need both the beginning and the 
end?
answer - because we might be able to write patterns that matches stuff that has different
lengths. We will get to those types of patterns.
end challenge

The motif str_something_all is also a general motif in the stringr package.
In general if we have a function that might return more than one result, we 
have a pair of functions. str_locate_all will find all the positions. 
str_locate() will find the first. If we are only interested in the first hit,
we can save some milliseconds by using str_locate() rather than str_locate_all().

In this case, only finding the first instance is about 5 times faster than 
finding all. The average for the two is 459 and 2277 respectively (on the computer
where these words are written). The unit is microseconds, so the win is small.

```{r}
falkner %>% 
  str_to_lower() %>% 
  str_locate("mother")
```



```{r}
library(tidyverse)
str_c()
#str_count()
#str_detect()
str_dup()
str_ends()
# str_extract()
# str_extract_all()
str_flatten()
str_glue()
str_glue_data()
str_interp()
str_length()
# str_locate()
# str_locate_all()
# str_match()
# str_match_all()
str_order()
str_pad()
# str_remove()
# str_remove_all()
# str_replace()
# str_replace_all()
str_replace_na()
str_sort()
str_split()
str_split_fixed()
str_squish()
str_starts()
str_sub()
str_subset()
# str_to_lower()
# str_to_sentence()
# str_to_title()
# str_to_upper()
str_trim()
str_trunc()
# str_view()
# str_view_all()
str_which()
str_wrap()
```

{% include links.md %}